Conference Paper Title*\\ {\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and should not be used} \thanks{Identify applicable funding agency here. If none, delete this.}

### 2023-05-20 LARS_paper/paper.tex
finish development (introduction)

According to the concepts presented, a small quadruped was designed and developed for research and educational purposes.(figuras \ref{fig:terrenos}, \ref{fig:caramel_body} e \ref{fig:moving_body}).
During the prototyping phase, the hardware architecture and physics structure was designed. Following that, the structure of the robot was printed and integrated with the hardware components. The robot's actuators are servo motors of the model dynamixel MX-28, and the central processing is composed of a Raspberry PI 4. Furthermore, the platform contains an inertial sensor, model MPU6050. All the software was developed with the Robot Operating System 2 Humble (ROS2) \cite{ROS2Humble}.

@@

The gait developed to the robot follows the sequence of phases demonstrated in figure \ref{fig:trot_pattern}, in which the blank rectangles represent the  swing phase and the gray rectangles represent the stance phase. During the stance phase, the legs are on the ground and push the robot in the desired direction.  In the swing phase, the legs are raised to move the legs to the next point of support. This is the main gait of the robot, called a trot, based on the concept of periodic and symmetrical gaits. The periodic gaits are characterized by the continuous repetition of the same movements in the same instants. While symmetric is a characteristic of gaits that move a pair of legs together, leaving and returning to the ground in a synchronized way. Although the structure of the robot makes it possible to perform other types of gaits, in this work, only the trot was considered. Still,  in order to reduce the complexity of locomotion control, a discontinuous gait was adopted, that is, the robot's body moves only when all the legs are on the ground, performing an intermittent movement. Still in figure \ref{fig:trot_pattern}, it is possible to realize that always the same pair of diagonal legs move at the same instant. Between two consecutive phases of swing, there is a moment when all legs are in stance, which is when the robot body is displaced in the desired direction of locomotion.

@@

O subsistema de controle do robô é composto por dois elementos principais: o controlador de angulação do corpo do robô e o planejador de marcha. O controle de angulação do corpo do robô é responsável por controlar o ângulo de \textit{pitch} (rotação em $y$) e o de \textit{roll} (rotação em $x$), de modo a manter o robô em $0^{\circ}$ a todo momento. O planejador de marcha é o responsável por controlar cada pata do robô e, por consequência, o corpo.  Ele define onde cada pata deve pisar e para onde o corpo deve se mover, com base nas etapas de \textit{stance} e \textit{swing}. O planejador de marcha utiliza um planejador de trajetória para calcular a trajetória que a pata deve realizar no espaço, de forma a mover o robô na direção e velocidade desejadas. Os pontos desta trajetória são convertidos nos ângulos de juntas através do modelo cinemático. O modelo cinemático do robô é responsável por mapear a posição tridimensional de cada pata e obter a angulação das juntas que a compõe.  Além disso, o planejador de trajetória também considera o esforço de controle enviado pelos controladores de angulação.

O planejador de marcha envia os pontos em que cada pata devem estar a uma frequência de $\SI{50}{\hertz}$. Essa frequência foi definida por ser a maior que o sistema de processamento do robô foi capaz de suportar e por respeitar o teorema Nyquist, que indica que, nesse caso, a frequência deve ser maior do que $\SI{4}{\hertz}$ --- já que o período do passo é $0.5 s$.

>>>>

According to the concepts presented, a small quadruped was designed and developed for research and educational purposes.(figuras \ref{fig:terrenos}, \ref{fig:caramel_body} e \ref{fig:moving_body}).
During the prototyping phase, the hardware architecture and physics structure was designed. Following that, the structure of the robot was printed and integrated with the hardware components. The robot's actuators are servo motors of the model dynamixel MX-28, and the central processing is composed of a Raspberry PI 4. Furthermore, it contains an inertial sensor, model MPU6050. All the software was developed with the Robot Operating System 2 Humble (ROS2) \cite{ROS2Humble}.

@@

The gait developed to the robot follows the sequence of phases demonstrated in figure \ref{fig:trot_pattern}, in which the blank rectangles represent the  swing phase and the gray rectangles represent the stance phase. During the stance phase, the legs are on the ground and push the robot in the desired direction.  In the swing phase, the legs are raised to move the legs to the next point of support. This is the main gait of the robot, called a trot, based on the concept of periodic and symmetrical gaits. The periodic gaits are characterized by the continuous repetition of the same movements in the same instants. While the symmetry is a characteristic of gaits that move a pair of legs together, leaving and returning to the ground in a synchronized way. Although the structure of the robot makes it possible to perform other types of gaits, in this work, only the trot was considered. Still,  in order to reduce the complexity of locomotion control, a discontinuous gait was adopted, that is, the robot's body moves only when all the legs are on the ground, performing an intermittent movement. In figure \ref{fig:trot_pattern}, it is possible to realize that always the same pair of diagonal legs move at the same instant. Between two consecutive phases of swing, there is a moment when all legs are in stance, which is when the robot body is displaced in the desired direction of locomotion.

@@

The control subsystem of the robot is composed of two main components: the rotation control and the gait planner. The rotation control of the body is responsible for controlling the pitch and angles of the robot, in order to keep it at 0 degrees at all times. The gait planner is responsible for controlling each leg and, consequently, the robot's body. It defines where the foot will step and where the body should move based on the swing and stance steps. The gait planner uses a trajectory planner to calculate the trajectory the foot should take in space in order to move the robot in the desired direction and speed. The points of this trajectory are converted into the angles of the joints through the kinematic model. The kinematic model of the robot is responsible for mapping the three-dimensional position of each leg and obtaining the joints' angles that compose it.  In addition, the trajectory planner also considers the control effort sent by the rotation controller.

The gait planner sends the points at which each foot must be at a frequency of $\SI{50}{\hertz}$. This frequency was defined as being the highest that the robot processing system was able to support and by respecting the Nyquist theorem, which indicates that in this case, the frequency should be greater than $\SI{4}{\hertz}$ -- since the step period is $0.5 s$.
